Partridge: An effective system for the automatic classification of the types of academic papers

### 2013-05-24 doc/bcspaper/article.tex
Maria latest edits

\usepackage{subcaption}
\usepackage{soul}
\captionsetup{compatibility=false}

@@

\abstract{Partridge is a system that enables intelligent search for
  academic papers by allowing users to query particular terms within
  sentences designating a particular core scientific concept (e.g. Hypotheses, Results, etc). 
  The system also automatically classifies papers according to
  article types (e.g. Review, Case Study). Here we focus on the latter
  aspect of the system. For each paper, Partridge automatically
  extracts the full paper content from PDF files, converts it to XML,
  determines sentence boundaries, automatically labels the sentences
  with core scientific concepts, and then uses a random forest model to
  classify the paper type. We show that the type of a paper can be
  reliably predicted by a model which analyses the distribution of
  core scientific concepts within the sentences of the paper. We
  discuss the appropriateness of many of the existing paper types used
  by major journals, and their corresponding distributions. Partridge
  is online and available for use, includes a browser-friendly
  bookmarklet for new paper submission, and demonstrates a range of
  possibilities for more intelligent search in the scientific
  literature. The Partidge instance can be used at
  \url{http://farnsworth.papro.org.uk}, and further information about
  the project can be found at \url{http://papro.org.uk}}

@@

This model of paper type is currently in use in our Partridge system, which has
been created as an intelligent full-text search platform for scientific papers.
Partridge makes use of automatically derived CoreSC sentence labels and
automatically derived paper types, to allow deeper information queries.  We
discuss the reliability of this model of paper types and the insights that have
been gained for the authorship of papers.

@@

\subsection*{Paper Processing and classification}
%Discuss in more detail how Partridge processes papers (PDFX, sentence splitting) 
In order to obtain automatic labeling of sentences in a paper with
CoreSC concepts we first needed to convert the papers to a format that
the the CoreSC classifier, SAPIENTA can analyse.  Currently, SAPIENTA
supports the SciXML and PubMed Central DTDs.  Papers in PDF format are
first converted to XML using PDFX %\cite{PDFX} - no paper known yet
, a free service hosted
by the University of Manchester (\url{http://pdfx.cs.man.ac.uk/}).
PDFX also provides output that already has some metadata, such as
title, author, and abstract, associated with it.  SAPIENTA allocates
each sentence within a paper a separate CoreSC label.  However, the
sentence boundaries must be detected before the document is passed
into SAPIENTA for annotation.  The sentence splitter uses a regular
expression to split the sentences within the document and then adds
the necessary \verb|<s>| tags to the markup to indicate the location
of each sentence.  Once the paper has been split, SAPIENTA is used to
annotate each sentence with a relevant CoreSC label.  These labels are
assigned using a conditional random fields model based upon a number
of features such as the sentence's location within the paper and pairs
and triplets of words found consecutively within the sentence
\cite{Liakata2012}.  The SAPIENTA system has been trained on a corpus
of 265 chemistry and biochemistry papers and we have done no domain
adaptation for the papers in PLoS.

@@

Random forest learning\cite{Breiman2001} was chosen for the construction of the
paper type classifier. This is because random forests are a fast, accurate and
widely accepted learning technology, and the underlying decision trees can be
inspected to understand some of the reasoning behind the predictions made by
the final paper type classifiers. The feature set consisted of the percentage
composition of each of the 11 respective CoreSC labels assigned to the
sentences in each paper i.e. (the percentage of {\em Background} sentences, the
percentage of {\em Hypothesis} sentences, etc.). The random forest learning was
conducted using the Orange data mining library for Python \cite{Curk2005}.  The
parameters for the random forest learner were as the defaults, except with
min\_subsets=5 and same\_majority\_pruning=true.  We used 10-fold cross
validation to estimate the precision, recall and F-measure.

@@

The results of the random forest learning as recall, precision and F-measure
for each paper type, averaged over a 10-fold cross validation are shown in Table
\ref{tab:recallPrecision}. Paper types {\em Research}, {\em Review} and {\em
Correspondence} are the most accurate classes, and {\em Viewpoint} is the most
difficult paper type to predict.

@@

A confusion matrix for the paper types in given in Table \ref{tab:confusion}.
Research papers are classified correctly most frequently, but are sometimes
confused with Case Study. However, Case Study papers are often predicted to be
Research or Review. There are fewer Case Study papers for the learning
algorithm to train on, and many research papers, so we would expect this
outcome. We might also expect that Research and Case Study would share similar
paper structures.  We initially expected the classes Essay, Opinion,
Perspective and Viewpoint to be confused, as the four labels all indicate a
paper containing an author's personal thoughts on an issue, rather than
traditional experiment-driven science. Opinion is confused with Perspective and
Viewpoint. Perspective is confused with almost all classes, except Research.
Viewpoint is a small class and mostly confused with Opinion and Perspective.
From this it would seem that there is a clear grouping of paper types which are
very similar, and could better be labelled as a single type: \{Opinion,
Perspective, Viewpoint\}. However, Essay appears to be different.

>>>>

\usepackage{subcaption}

\captionsetup{compatibility=false}

@@

\abstract{Partridge is a system that enables intelligent search for
  academic papers by allowing users to query terms within
  sentences designating a particular core scientific concept (e.g. Hypothesis, Result, etc). 
  The system also automatically classifies papers according to
  article types (e.g. Review, Case Study). Here we focus on the latter
  aspect of the system. For each paper, Partridge automatically
  extracts the full paper content from PDF files, converts it to XML,
  determines sentence boundaries, automatically labels the sentences
  with core scientific concepts, and then uses a random forest model to
  classify the paper type. We show that the type of a paper can be
  reliably predicted by a model which analyses the distribution of
  core scientific concepts within the sentences of the paper. We
  discuss the appropriateness of many of the existing paper types used
  by major journals, and their corresponding distributions. Partridge
  is online and available for use, includes a browser-friendly
  bookmarklet for new paper submission, and demonstrates a range of
  possibilities for more intelligent search in the scientific
  literature. The Partidge instance can be used at
  \url{http://farnsworth.papro.org.uk}, and further information about
  the project can be found at \url{http://papro.org.uk}}

@@

This model of paper type is currently in use in our Partridge system, which has
been created as an intelligent full-text search platform for scientific papers.
Partridge (which currently holds XXX papers and is constantly expanding) 
makes use of automatically derived CoreSC sentence labels and
automatically derived paper types, to allow deeper information queries.  We
discuss the reliability of this model of paper types and the insights that have
been gained for the authorship of papers.

@@

\subsection*{Paper Processing and classification}
%Discuss in more detail how Partridge processes papers (PDFX, sentence splitting) 
In order to obtain automatic labeling of sentences in a paper with
CoreSC concepts we first needed to convert the papers to a format that
the the CoreSC classifier, SAPIENTA can analyse.  Currently, SAPIENTA
supports the SciXML and PubMed Central DTDs.  Papers in PDF format are
first converted to XML using PDFX %\cite{PDFX} - no paper known yet
, a free service hosted by the University of Manchester
(\url{http://pdfx.cs.man.ac.uk/}). Once the paper has been split into
sentences, SAPIENTA is used to annotate each sentence with a relevant
CoreSC label. These labels are assigned using a conditional random
fields model based upon a number of features such as the sentence's
location within the paper and pairs and triplets of words found
consecutively within the sentence \cite{Liakata2012}.  The SAPIENTA
system has been trained on a corpus of 265 chemistry and biochemistry
papers and we have done no domain adaptation for the papers in PLoS.

@@

Machine learning to build a classifer of paper type was conducted using random
forest learning \cite{Breiman2001}.  A random forest learning algorithm was
chosen because this is a fast, accurate and widely accepted learning
technology, and because the underlying decision trees can be used to describe
and inspect some of the reasoning behind the predictions made by the final
paper type classifiers.  The feature set consisted of the percentage
composition of the CoreSC labels given to the sentences in each paper i.e. (the
percentage of {\em Background} sentences, the percentage of {\em Hypothesis}
sentences, etc.).  Thus there are 11 features corresponding to the 11 CoreSC
labels.  The random forest learning was conducted using the Orange data mining
library for Python \cite{Curk2005}.  The parameters for the random forest
learner were as the defaults, except with min\_subsets=5 and same\_majority\_pruning=true.  
We used 10-fold cross validation to estimate the precision,
recall and F-measure.

@@

We report the results of the random forest learning as recall, precision and F-measure for each paper type, averaged over a 10-fold cross validation in Table \ref{tab:recallPrecision}. Paper types {\em Research}, {\em Review} and {\em Correspondence} are the most accurate classes, and {\em Viewpoint} is the most difficult paper type to predict.

@@

A confusion matrix for the paper types in given in Table \ref{tab:confusion}. Research papers are usually predicted to belong to the Research class, and are sometimes confused with Case Study. However, Case Study papers are often predicted to be Research or Review. There are fewer Case Study papers for the learning algorithm to train on, and many research papers, so we would expect this outcome. We might also expect that Research and Case Study would share similar paper structures. 
We initially expected the classes Essay, Opinion, Perspective and Viewpoint to be confused, as the four labels all indicate a paper containing an author's personal thoughts on an issue, rather than traditional experiment-driven science. Opinion is confused with Perspective and Viewpoint. Perspective is confused with almost all classes, except Research. Viewpoint is a small class and mostly confused with Opinion and Perspective. From this it would seem that there is a clear grouping of paper types which are very similar, and could better be labelled as a single type: \{Opinion, Perspective, Viewpoint\}. However, Essay appears to be different.
